Segmentation of Electron Microscopy Images for Connectomics

### 2018-05-01 papers/quals2018/0_introduction.tex
quals: major updates to structure

\section{Introduction}

The field of connectomics is concerned with the reconstruction of the wiring diagram of the mammalian brain.
In order to achieve this ambitious goal, we need images at nanometer resolution to see the structure of the neurons and the synaptic connections between them.
Recent advances in electron microscopy acquisition techniques enable a throughput of a terabyte of image data every hour~\cite{richard2016imaging}.
Manual reconstruction of this image data is simply infeasible and requires automatic segmentation methods.
These automatic methods label every voxel of the image with a 64-bit integer value where two voxels have the same label only if they belong to the same neuron.
As the field progresses, our goal is to reconstruct a cubic millimeter of brain which will result in over 18 petabytes of segmentation data alone~\cite{suissa2016automatic}.

Three major challenges arise as the image datasets scale to a petabyte and more. 
First, the segmentation methods must be accurate enough to extract the proper connections between neurons.
In particular, the dendritic spines \TODO{check,CITE} which contain the postsynaptic density can be only \TODO{XX} nanometers thick.
Second, automatic segmentation must be fast -- ideally matching the throughput of the electron microscope~\cite{haehn2017scalable}.
The current state-of-the-art automatic reconstruction pipelines are either too slow for such large datasets or too inaccurate at scale.
Third, the label volumes require $8\times$ the number of bytes as the already massive image volumes.
Storing such data is expensive and transferring the data is slow. 
To cut costs and delays, we need compression methods to reduce data sizes.

>>>>

% !TEX root =  paper.tex
\section{Introduction}

One of the critical components of connectomics---the field concerned with reconstructing the wiring diagram of the brain at nanometer resolution---is automatic segmentation of electron microscopy (EM) images of the brain.
With recent advances in EM acquisition techniques, neuroscientists can now generate a terabyte of image data every hour~\cite{richard2016imaging}.
These images are typically anisotropic with $4$nm resolution in the $xy$ plane and \numrange{30}{40}nm between slices. 
At this resolution, we can see the axon terminals and dendritic spines, and the synaptic connections between them.
Accurate reconstruction of every neuron paired with synapse detection will allow us to create a true graphical representation of the brain and further our understanding of the underlying circuitry.
These segmentation, or label, volumes assign a 32- or 64-bit integer label to every voxel where voxels have the same label only if they belong to the same neuron.

An ambitious manual reconstruction effort in the 1980s resulted in the first complete connectome of any animal, the \textit{Caenorhabditis elegans} worm~\cite{white1986structure}. 
This species has 302 neurons and manual labeling required several years.
More recent studies focus on \textit{Drosophila} flies~\cite{jovanic2016competitive,takemura2015synaptic},  rodents~\cite{richard2016imaging}, and even humans~\cite{sporns2005human}. 
With an EM throughput of one terabyte per hour, neuroscientists can image a cubic millimeter of data (two petabytes) in six months~\cite{suissa2016automatic}. 
At this scale, the manual reconstruction techniques of the 80s are simply infeasible.
Thus, researchers rely on deep learning methods to segment these massive datasets.

Automatic reconstruction produces label volumes that, uncompressed, are much larger in size than the already massive image datasets.
In our paper published in \textit{MICCAI 2017}, we explore the effectiveness of existing general-purpose, image, and video compression techniques on the segmentation datasets.
We find that these existing techniques fail to adequately exploit the typical properties of our label volumes.
Thus, we propose \textit{Compresso}, a new compression algorithm specifically tailored for large label volumes that far exceeds existing strategies.

Automatic reconstruction techniques need to be fast, scalable, and accurate.
Ideally, image acquisition is the bottleneck of the connectomics pipeline so reconstruction occurs at a rate of one terabyte per hour~\cite{haehn2017scalable}.
We can achieve this throughput with parallelization among several GPUs with existing techniques~\cite{funke2017deep,parag2017anisotropic}.
However, these automatic methods typically only rely on local context for decision making and are agnostic about the underlying biological systems.
Thus, they often make errors at scale and currently require human proofreading or other error correction techniques~\cite{haehn2017guided,error_correction_using_CNN}.

We propose a novel region merging framework that takes as input an oversegmentation of an EM volume (under review, \textit{ECCV 2018}).
Our method imposes both local and global biological constraints onto the output segmentation to more closely match the underlying structure of neuronal processes.
Additionally, our method is independent of image resolution and acquisition parameters, enabling its application to isotropic and anisotropic data without retraining.
Images generated by electron microscopes often differ in appearance because of variations in staining techniques~\cite{briggman2012volume}.
By removing the dependence of our algorithm on the input images, we greatly reduce the need for additional costly ground truth data for each new stack of images.

### 2018-05-01 papers/quals2018/1_related_work.tex
quals: major updates to structure

\section{Related Work}

\begin{figure*}[t!]
	\centering
	\includegraphics[width=0.75\linewidth]{./figures/teaser_v4.png}
	\caption{Our framework uses both geometric and topological biological constraints for region merging. We (a) generate a simplified graph from the skeleton representation of the input segmentation, (b) train a classifier to learn the edge weight based on the shape of the segment connection, and (c) perform graph partitioning with acyclic constraints.}
	\label{fig:teaser_pipeline}
\end{figure*}

A significant amount of connectomics research considers the problem of extracting segmentation information at the voxel level of EM images.
First, intermediate representations like boundary, affinity or binary segmentation maps are generated from the voxels.
Random forests with hand-designed features~\cite{kaynig2015large}, or 2D and 3D convolutional networks produce boundary probabilities~\cite{seymour2016rhoananet,ronneberger2015u,bogovic2013learned,ciresan2012deep,jain2010boundary,amelio_segmentation}.
Often, the affinity between each voxel and its six neighbors are used~\cite{lee2015recursive,parag2017anisotropic,lee2017superhuman,cciccek20163d,turaga2010convolutional}. 
The 3D U-Net architecture has become popular~\cite{cciccek20163d} and the MALIS cost function is specifically designed to re-weight affinity predictions by their contribution to the segmentation error~\cite{briggman2009maximin}.
More recently, flood-filling networks~\cite{januszewski2016flood} produce binary segmentations from raw pixels with a recurrent convolutional network at a high computational cost.
Orthogonal to the representation, model averaging~\cite{zeng2017deepem3d} and data augmentation~\cite{lee2017superhuman} methods can further improve the performance, where Lee et al.~\cite{lee2017superhuman} surpass the estimated human accuracy on the SNEMI3D dataset.

@@

Additional research builds on top of these region-based methods to correct errors in the segmentation. This can be done either using human proofreading~\cite{haehn2017guided,haehn2014design,mojo2} or automatically~\cite{rolnick2017morphological,error_correction_using_CNN}. In both cases, available methods are pixel-based and do not include global biological constraints into the decision making process. Our method can be used as input to any existing error correction framework.

>>>>

% !TEX root =  paper.tex
\section{Related Work}

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=0.45\linewidth]{./figures/encoding_diagram_opt.pdf}
		\hspace{0.08\linewidth}
		\includegraphics[width=0.45\linewidth]{./figures/window_values.pdf}
	\end{center}	
	\caption{Compresso works by dividing up the segmentation data into small blocks. On the left we show the intersection of three segments in a $4\times4\times1$ window. We extract a boundary map from this segmentation to transform the window into an integral value between $0$ and $2^{16} - 1$. The location $i$ requires some additional bookkeeping. On the right are the $100$ most common $8 \times 8 \times 1$ windows accounting for $\sim82\%$ of the volume on a representative dataset.}
	\label{fig:compression}
\end{figure*}

General-purpose compression schemes~\cite{collet2016smaller,deutsch1996zlib,google2016brotli,lehmann2016liblzf,oberhumer2005lzo,pavlov2007lzma,seward1998bzip2,vandevenne2016zopfli,welch1984technique,ziv1978compression} are not optimized for this data.
With Compresso, an algorithm published in MICCAI 2017, we exploit the typical characteristics of label volumes such as large invariant regions without natural relationship between label values. 
These properties render 2D image compression schemes inadequate since they rely on frequency reduction (using e.g., wavelet or discrete cosine transform) and value prediction of pixels based on local context (differential pulse-code modulation)~\cite{roelofs1999png,skodras2001jpeg}. 
Color space optimization strategies in video codecs~\cite{aimar2005x264} also have no effect on label volumes, even though the spatial properties of a segmentation stack (\textit{z}-axis) are similar to the temporal properties of video data (time-axis). 
A compression scheme designed specifically for label volumes is part of the visualization software Neuroglancer~\cite{google2016compressed}. 
This method exploits segmentation homogeneity by creating small blocks with $N$ labels and reducing local entropy to $\log_2{N}$ per pixel. 
Lookup tables then decode the values $[0,N)$ to the original 64-bit labels.

A significant amount of connectomics research considers the problem of extracting segmentation information at the voxel level of EM images.
First, intermediate representations like boundary, affinity or binary segmentation maps are generated from the voxels.
Random forests with hand-designed features~\cite{kaynig2015large}, or 2D and 3D convolutional networks produce boundary probabilities~\cite{bogovic2013learned,ciresan2012deep,jain2010boundary,seymour2016rhoananet,ronneberger2015u,amelio_segmentation}.
Often, the affinity between each voxel and its six neighbors are used~\cite{cciccek20163d,lee2015recursive,lee2017superhuman,parag2017anisotropic,turaga2010convolutional}. 
The 3D U-Net architecture has become popular~\cite{cciccek20163d} and the MALIS cost function is specifically designed to re-weight affinity predictions by their contribution to the segmentation error~\cite{briggman2009maximin}.
More recently, flood-filling networks~\cite{januszewski2016flood} produce binary segmentations from raw pixels with a recurrent convolutional network at a high computational cost.
Orthogonal to the representation, model averaging~\cite{zeng2017deepem3d} and data augmentation~\cite{lee2017superhuman} methods can further improve the performance, where Lee et al.~\cite{lee2017superhuman} surpass the estimated human accuracy on the SNEMI3D dataset.

@@

Additional research builds on top of these region-based methods to correct errors in the segmentation. This can be done either using human proofreading~\cite{haehn2017guided,haehn2014design,mojo2} or automatically~\cite{rolnick2017morphological,error_correction_using_CNN}. 
In both cases, available methods are pixel-based and do not include global biological constraints into the decision making process. 
Our method can take as input any existing segmentation pipeline.

### 2018-05-01 papers/quals2018/2_method.tex
quals: added method and evaluation from eccv submission
quals: major updates to structure

% !TEX root =  paper.tex

\section{Method}
%\begin{figure*}[t!]
%	\begin{center}
%		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut_input.png}}
%		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/pre-multicut.png}}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/multicut-graph.png}}
% 		\subfloat[]{\includegraphics[width=0.24\linewidth]{./figures/methods/post-multicut.png}}
% 	\end{center}
% 	\caption{Outline of our approach, from left to right: result of the pixel-based segmentation and agglomeration algorithm; segments of several neurons from the initial segmentation; extracted skeletonized network of those segments; improved 3D reconstruction of the selected segments after graph construction and partitioning with constraints.}
% 	\label{fig:overview}
% \end{figure*}

%There are two types of errors that can occur in connectomics segmentations: \textit{split errors} and \textit{merge errors}. 
%In a split error, two segments should have been merged into one neuronal process (Fig.~\ref{fig:improved-reconstruction}). 
%In a merge error, one segment corresponds to more than one neuron.
%Generally, it is much more difficult to correct merge errors than to correct split errors, as the space of possible split proposals grows quickly~\cite{parag2015properties}.
%Thus, most reconstruction approaches are tuned towards over-segmentation with many more split than merge errors. 
%The input of our method is the over-segmentation of EM image volumes from state-of-the-art pipelines.

Our method is illustrated in Fig.~\ref{fig:teaser_pipeline}.
From the input segmentation we generate a graph $G$ with nodes $N$ and edges $E$ with weights $w_e$. 
The nodes correspond to labeled segments from the input data with edges between merge candidates.
We propose the following three steps to formulate and then partition the graph while obeying constraints from the underlying biology.
First, we only consider merging segments based on a skeletonized representation of the input segmentation (Fig.~\ref{fig:teaser_pipeline}a).
This enables us to reduce the number of nodes in the graph based on prior knowledge on the shape of neuronal processes.
Second, we train a convolutional neural network that learns biological constraints based on the shapes of the input segmentation (Fig.~\ref{fig:teaser_pipeline}b).
This network generates probabilities that segments belong to the same neuron based only on the segmentations. 
An example of one such learned constraint is that neurons have small turning radii (Fig.~\ref{fig:turn-radii}).
Third, we partition the graph using a lifted multicut formulation with additional acyclic constraints to enforce global biological constraints (Fig.~\ref{fig:teaser_pipeline}c).
The lifted multicut solution is globally consistent which we then augment to produce a tree-structured graph (i.e., one with no cycles).

\subsection{Skeleton-Based Graph Generation}
%\subsubsection{Node Generation}
\label{sec:skeletonization}
Most region merging methods create a region graph by removing small-sized segments and linking each pair of adjacent segments, which can still lead to a large graph size due to the irregular shape of neural structures.
We use a skeleton representation of the segmentation to reduce the graph size with the geometric constraints on the connectivity of two adjacent segments to prune edges.
% a bit trivial, as everyone is doing this. maybe no need to mention
% \subsubsection{Node Pruning}
% %The simplest node generation strategy creates one node for every unique segment label in the input volume. 
% Some labels in the volume correspond to very small structures that are likely the result of segmentation errors, typically in regions with noisy raw image data. 
% It is difficult to extract useful shape features from these segments because of their small, often random, shape. 
% We prune these nodes from the graph by removing all segments with fewer than a empirical threshold $t_{seg} = 20,000$ voxels. 
% This removes on average \TODO{XX}\% of the segments in our  datasets (Sec.~\ref{sec:dataset}), leaving us with around $\TODO{XX}$ nodes per dataset (\TODO{X nodes per micrometer}). 
% Despite the large number of segments, these regions only take up \TODO{XX}\% of the total volume on average.

% \subsubsection{Edge Pruning}

\begin{figure}[t]
	\centering
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/skeleton1.png}		
	\end{minipage}
	\hfill
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/skeleton2.png}		
	\end{minipage}
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/skeleton3.png}		
	\end{minipage}
	\hfill
	\begin{minipage}{0.45\linewidth}
		\includegraphics[width=\linewidth]{./figures/skeleton4.png}		
	\end{minipage}
	\caption{Example skeletons (in black) extracted from segments using a variant of the TEASER algorithm~\cite{sato2000teasar}. These skeletons not only capture the shape of the segmentation, but also provide endpoints useful for region merging proposals.}
	\label{fig:skeletonization}
\end{figure}

%A typical approach for generating edges produces one between all adjacent segments. 
%Two segments $l_1$ and $l_2$ are considered adjacent if there is a pair of adjacent voxels with one labeled $l_1$ and the other labeled $l_2$.
%For example, pixel-based agglomeration methods such as mean agglomeration~\cite{lee2017superhuman} or waterz~\cite{funke2017deep} consider all pairs of adjacent segments for merging.
%However, this method produces too many edges in the graph for graph-based optimization approaches. 
Our key observation is that if two segments belong to the same neuronal process, their skeleton end points should satisfy certain geometric constraints.
To extract the skeleton from each segment, we use a variant~\cite{zhao2014automatic} of the TEASER algorithm~\cite{sato2000teasar}. 
This skeletonization algorithm repeatedly uses Dijkstra's algorithm to find the farthest voxel from a seed location. 
Since this algorithm is non-linear in the number of voxels, we downsample the datasets so that there are voxel samples every $\SI{30}{\nano\meter}$ in each dimension.
Empirically, this reduced the running time for skeletonization by $30\times$ with minimal reduction in skeleton accuracy (${\sim}5\%$ fewer branches). 
Fig.~\ref{fig:skeletonization} shows four examples of extracted skeletons (in black). 
These skeletons consist of a sequence of \textit{joints}, locations that are locally a maximum distance from the segment boundary, with line segments connecting successive joints. 
We refer to joints that have only one connected neighbor as \textit{endpoints}. 
We find that approximately 70\% of the segments that are erroneously split have nearby endpoints (Fig.~\ref{fig:merge_candidates}).

Two segments, $s_1$ and $s_2$, receive a corresponding edge if the two following conditions hold.
First, endpoints in either $s_1$ or $s_2$ are within $t_{low}$ nm of any voxel in the other segment.
Second, there are endpoints in $s_1$ and in $s_2$ that are within $t_{high}$ nm of each other.
We store the midpoints between the two endpoints as the center of the potential merge in the set $\mathbb{S}_c$. 
This algorithm produces a set of segments to consider for merging. 
Only these pairs have a corresponding edge in the constructed graph.
We provide an empirical analysis of these parameters on the structure of the graph in the supplemental materials.

\begin{figure}[t]
	\begin{minipage}{0.4\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_error.png}
	\end{minipage}
	\hfill
	\begin{minipage}{0.55\linewidth}
		\includegraphics[width=\linewidth]{./figures/constraint_success.png}
	\end{minipage}
	\caption{Geometric constraints for region merging. We show two region merging proposals. On the left, the segments do not belong to the same neuron, as evidenced by the sharp turning radius indicated by the arrow; while on the right, the segments should be merged due to the continuity of the 3D shape. Instead of using handcrafted geometric features, we train a convolutional neural networks to automatically learn them from the ground truth labels.}
	\label{fig:turn-radii}
\end{figure}
\subsection{Learning-based Edge Weight}
\label{sec:edge-weights}
We assign an edge weight $w_e$ to each edge corresponding to the probability that two nodes belong to the same neuronal process.
We train a 3D CNN model to learn these geometric constraints for valid connection between two segments from labeled volume.
\\~\\
\noindent\textbf{Edge Probability}
To predict the probabilities that two segments belong to the same neuron, we train a feed-forward convolutional network with three \textit{VGG-style} convolution blocks~\cite{chatfield2014return} and two fully connected layers before the final sigmoid activation.

For the input of the network, we extract a cubic region of interest (ROI) around each endpoint $e$ in $\mathbb{S}_c$ as input to the CNN. 
The CNN receives three input channels for every voxel in the ROI around segments $l_1$ and $l_2$. 
The input in all of the channels is in the set $\{-0.5, 0.5\}$. 
The first channel is $0.5$ only if the corresponding voxel has label $l_1$. 
The second channel is $0.5$ only if the corresponding voxel has label $l_2$. 
The third channel is $0.5$ if the corresponding voxel is either $l_1$ or $l_2$.
We do not use the raw EM image information to avoid the need to retrain the network on datasets that have been stained differently or imaged at different resolution. 
This reduces the need for generating costly manually-labeled ground truth.

%The first two max pooling layers are anisotropic with pooling only in the $x$ and $y$ dimensions. 
%The output of this final pooling step is flattened into a 1D vector that is input into . 
%The final layer produces probabilities with a sigmoid activation function~\cite{funahashi1989approximate}. 
%All of the other activation functions are .

% covered in the experiment
%For training we use a stochastic gradient descent optimizer with Nesterov's accelerated gradient~\cite{nesterov1983method}. 
%We employ dropouts of $0.2$ after every pooling layer and the first dense layer, and a dropout of $0.5$ after the final dense layer to prevent overfitting. 
%We discuss all other network parameters in Sec.~\ref{sec:network-parameters}.

\begin{figure}[t]
	\centering
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error1.png}		
	\end{minipage}
	\hfill
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error2.png}				
	\end{minipage}
	\hfill
	\begin{minipage}{0.32\linewidth}
		\includegraphics[width=\linewidth]{./figures/split_error3.png}
	\end{minipage}
	\caption{Three erroneously split segments.}
	\label{fig:merge_candidates}
\end{figure}
\subsection{Optimization-Based Graph Partition}
\label{sec:optimization}
After graph construction, we segment the graph in a globally consistent manner while enforcing topological constraints on the output.
\\~\\
%\subsubsection{Lifted Multicut}
\noindent\textbf{Lifted Multicut}
After constructing the graph we seek to partition it into labels where every label corresponds to a neuronal process. 
We formulate this graph partitioning problem as a multicut problem.
There are two primary benefits to using a multicut formulation. 
First, the number of segments in the final graph is not predetermined but depends on the input. 
Second, this minimization produces globally consistent solutions (i.e., a boundary remains only if the two corresponding nodes belong to unique segments)~\cite{keuper2015efficient}.

We apply the algorithms of Keuper et al.~\cite{keuper2015efficient} to produce a feasible solution to the multicut problem using greedy additive edge contraction.
Following their example, we employ the generalized lifted multicut formulation.
Traditional multicut solutions only consider the probabilities that two adjacent nodes belong to the same segment. 
In the lifted extension to the problem, we can penalize non-adjacent nodes that belong to different segments. 
These penalties between non-adjacent nodes are called lifted edges. 
Ideally these lifted weights represent the probability that two nodes belong to the same neuron.
However, determining such probabilities is computationally expensive.
We approximate these probabilities by finding the maximal probable path between any two nodes using Dijkstra's algorithm~\cite{keuper2015efficient}.
This is an underestimate of the probability that two nodes belong to the same neuron since it does not consider all possible paths.
Since our graphs are sufficiently small, we can generate lifted edges between all pairs of nodes. 
\\~\\
\noindent\textbf{Edge Weight} 
We need to convert the probabilities into the following weighting scheme to solve the multicut problem with this heuristic~\cite{keuper2015efficient,andres2011probabilistic}.
Given the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e$ is defined as
%p_e &=\mbox{CNN}(x_1, x_2) \\
\begin{align}
w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}},
\end{align}
where $\beta$ is a tunable parameter that encourages over- or under-segmentation. 
Since, there are many more lifted edges than adjacent edges, we scale down the lifted weights proportionally to their total number  
~\cite{beier2017multicut}.
\\~\\
%\subsubsection{Topological Constraints}
\noindent\textbf{Topological Constraints}
By reformulating the segmentation problem as a graph partitioning one, we can enforce some global constraints on our result based on the underlying biology.
Traditional hierarchical clustering algorithms do not rely on such constraints but consider local decisions independently.
We enforce a global constraint that neurons are tree-structured and should not contain cycles. 
The multicut problems returns a series of ``collapsed'' edges between nodes that belong to the same neuron.
We iterate over these edges in order of the probability of merge generated by our CNN. 
We ``collapse'' an edge only if it does not create a cycle in the graph.

>>>>

% !TEX root =  paper.tex
\section{Biologically-Constrained Error Correction}

\begin{figure*}[t!]
	\centering
	\includegraphics[width=0.75\linewidth]{./figures/teaser_v4.png}
	\caption{Our framework uses both geometric and topological biological constraints for region merging. We (a) generate a simplified graph from the skeleton representation of the input segmentation, (b) train a classifier to learn the edge weight based on the shape of the segment connection, and (c) perform graph partitioning with acyclic constraints.}
	\label{fig:teaser_pipeline}
\end{figure*}

Our method is illustrated in Fig.~\ref{fig:teaser_pipeline}.
From the input segmentation we generate a graph $G$ with nodes $N$ and edges $E$ with weights $w_e$. 
The nodes correspond to labeled segments from the input data with edges between merge candidates.
We propose the following three steps to formulate and then partition the graph while obeying constraints from the underlying biology.
First, we only consider merging segments based on a skeletonized representation of the input segmentation (Fig.~\ref{fig:teaser_pipeline}a).
This enables us to reduce the number of edges in the graph based on prior knowledge on the shape of neuronal processes.
Second, we train a convolutional neural network that learns biological constraints based on the shapes of the input segmentation (Fig.~\ref{fig:teaser_pipeline}b).
This network generates probabilities that segments belong to the same neuron based only on the segmentations. 
An example of one such learned constraint is that neurons have small turning radii (Fig.~\ref{fig:turn-radii}).
Third, we partition the graph using a lifted multicut formulation with additional acyclic constraints to enforce global biological constraints (Fig.~\ref{fig:teaser_pipeline}c).
The lifted multicut solution is globally consistent which we then augment to produce a tree-structured graph.

\subsection{Skeleton-Based Graph Generation}
\label{sec:skeletonization}
Most region merging methods create a region graph by removing small-sized segments and linking each pair of adjacent segments, which can still lead to a large graph size due to the irregular shape of neural structures.
We use a skeleton representation of the segmentation to reduce the graph size with geometric constraints on the connectivity of two adjacent segments to prune edges.

Our key observation is that if two segments belong to the same neuronal process, their skeleton end points should satisfy certain geometric constraints.
To extract the skeleton from each segment, we use a variant~\cite{zhao2014automatic} of the TEASER algorithm~\cite{sato2000teasar}. 
Fig.~\ref{fig:skeletonization} shows four examples of extracted skeletons (in black). 
These skeletons consist of a sequence of \textit{joints}, locations that are locally a maximum distance from the segment boundary, with line segments connecting successive joints. 
We refer to joints that have only one connected neighbor as \textit{endpoints}. 
We find that approximately 70\% of the segments that are erroneously split have nearby endpoints (Fig.~\ref{fig:merge_candidates}).

Two segments, $s_1$ and $s_2$, receive a corresponding edge if the two following conditions hold.
First, endpoints in either $s_1$ or $s_2$ are within $t_{low}$ nm of any voxel in the other segment.
Second, there are endpoints in $s_1$ and in $s_2$ that are within $t_{high}$ nm of each other.
We store the midpoints between the two endpoints as the center of the potential merge in the set $\mathbb{S}_c$. 
This algorithm produces a set of segments to consider for merging. 
Only these pairs have a corresponding edge in the constructed graph.

\subsection{Learning-based Edge Weight}
\label{sec:edge-weights}

To predict the probabilities that two segments belong to the same neuron, we train a feed-forward convolutional network with three \textit{VGG-style} convolution blocks~\cite{chatfield2014return} and two fully connected layers before the final sigmoid activation. 
For the input of the network, we extract a cubic region of interest (ROI) around each endpoint $e$ in $\mathbb{S}_c$ as input to the CNN. 
There are three channels corresponding to voxels belonging to label one, label two, or either label.
We do not use the raw EM image information to avoid the need to retrain the network on datasets that have been stained differently or imaged at different resolution. 
This reduces the need for generating costly manually-labeled ground truth.

\subsection{Optimization-Based Graph Partition}

\label{sec:optimization}

After constructing the graph we seek to partition it into labels where every label corresponds to a neuronal process. 
We formulate this graph partitioning problem as a multicut problem.
There are two primary benefits to using a multicut formulation. 
First, the number of segments in the final graph is not predetermined but depends on the input. 
Second, this minimization produces globally consistent solutions (i.e., a boundary remains only if the two corresponding nodes belong to unique segments)~\cite{keuper2015efficient}.

We apply the algorithms of Keuper et al.~\cite{keuper2015efficient} to produce a feasible solution to the multicut problem using greedy additive edge contraction.
Following their example, we employ the generalized lifted multicut formulation.
Traditional multicut solutions only consider the probabilities that two adjacent nodes belong to the same segment. 
In the lifted extension to the problem, we can penalize non-adjacent nodes that belong to different segments. 
These penalties between non-adjacent nodes are called lifted edges. 
Ideally these lifted weights represent the probability that two nodes belong to the same neuron.
However, determining such probabilities is computationally expensive.
We approximate these probabilities by finding the maximal probable path between any two nodes using Dijkstra's algorithm~\cite{keuper2015efficient}.
This is an underestimate of the probability that two nodes belong to the same neuron since it does not consider all possible paths.
Since our graphs are sufficiently small, we can generate lifted edges between all pairs of nodes.

We need to convert the probabilities into the following weighting scheme to solve the multicut problem with this heuristic~\cite{andres2011probabilistic,keuper2015efficient}.
Given the probability that the nodes belong to the same neuron is $p_e$, the edge weight $w_e$ is defined as
%p_e &=\mbox{CNN}(x_1, x_2) \\
\begin{align}
w_e = \log{\frac{p_e}{1 - p_e}} + \log{\frac{1 - \beta}{\beta}},
\end{align}
where $\beta$ is a tunable parameter that encourages over- or under-segmentation. 
Since, there are many more lifted edges than adjacent edges, we scale down the lifted weights proportionally to their total number  
~\cite{beier2017multicut}.

By reformulating the segmentation problem as a graph partitioning one, we can enforce some global constraints on our result based on the underlying biology.
Traditional hierarchical clustering algorithms do not rely on such constraints but consider local decisions independently.
We enforce a global constraint that neurons are tree-structured and should not contain cycles. 
The multicut problems returns a series of ``collapsed'' edges between nodes that belong to the same neuron.
We iterate over these edges in order of the probability of merge generated by our CNN. 
We ``collapse'' an edge only if it does not create a cycle in the graph.
